---
title: "Sparcc Simulation 1"
author: "Anonymous"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
library(ggh4x)
library(scales)
library(ggthemes)
library(kableExtra)

```

## Simulation Setup

The goal of this program is to evaluate the performance of the semiparametric efficient estimator $\widehat{\pmb{\beta}}_{sp}$ using simulations in the following setting:

* $Z \sim \textrm{Bernoulli}(0.5)$,
* $X|Z \sim \textrm{beta}(\alpha_1 + \alpha_2 Z, \alpha_3 + \alpha_4 Z)$,
* $C|Z \sim \textrm{beta}(\gamma_1 + \gamma_2 Z, \gamma_3 + \gamma_4 Z)$,
* $Y|X,Z \sim N(\beta_0 + \beta_1X + \beta_2Z, \sigma^2)$.


In these simulations, we vary the censoring probability $q=\textrm{P}(X>C)$.

We also vary which how nuisance distributions are modeled: $X|Z$ (or $C|Z$) correctly specified as beta conditional on $Z$ with estimated parameters $\pmb{\alpha}, \pmb{\gamma}$, $X|Z$ (or $C|Z$) incorrectly specified as marginal beta, or both $X|Z$ and $C|Z$ nonparametrically estimated.

For each simulation, we generate a data set of size $n$ then compute the complete case estimator $\widehat{\pmb{\beta}}_{cc}$, the oracle estimator $\widehat{\pmb{\beta}}_{or}$, the parametric MLE $\widehat{\pmb{\beta}}_{ml}$, and the semiparametric efficient estimator $\widehat{\pmb{\beta}}_{sp}$.

```{r}

# true beta
B.true <- c(1, 10, 2, log(1))

# load simulation results from each of 10 clusters
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0(
            "sim1_data/sd",
            clust, ".csv")))
  })

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

colnames(sim.out) <- gsub(".B", ".", colnames(sim.out))
colnames(sim.out) <- gsub("B2", "trueB2", colnames(sim.out))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with(c("B", "V")),
               names_to = "method.param",
               values_to = "estimate") %>% 
  mutate(name = factor(str_sub(method.param, 1, 1)),
         method = factor(str_sub(method.param, 3, 4),
                         levels = c("sp", "ml", "cc", "or"),
                         labels = c("Semipar", "MLE", "CC", "Oracle")),
         nuis.x = factor(str_sub(str_sub(
           gsub("[.]", "", method.param), 4, -2), 1, 1),
           levels = c("", "1", "0",  "2"),
           labels = c("-", "Correct", "Incorrect", "Nonpar")),
         nuis.c = factor(str_sub(str_sub(
           gsub("[.]", "", method.param), 4, -2), 2, 2),
           levels = c("", "1", "0",  "2"),
           labels = c("-", "Correct", "Incorrect", "Nonpar")),
         param = factor(str_sub(method.param, -1, -1)),
         B.true = B.true[param]) %>% 
  dplyr::select(-c(method.param)) %>% 
  group_by(clust, n, q, name, method, nuis.x, nuis.c, param) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = estimate,
              id_cols = c(clust, n, q, method, nuis.x, nuis.c,
                          param, B.true, id)) %>% 
  mutate(B = ifelse(abs(B - B.true) > 10, NA, B),
         ci.lower = B - qnorm(0.975) * V,
         ci.upper = B + qnorm(0.975) * V,
         ci.cov = B.true >= ci.lower & B.true <= ci.upper)
             
```

## Simulation Results

#### Errors in Simulations

We first look at which simulation settings lead to errors.

```{r message = F}

# check for simulations with errors
sim.out.long %>% 
  filter(param == 1) %>% 
  group_by(q, n, method, nuis.x, nuis.c) %>% 
  summarize(n.error = sum(is.na(B)),
            prop.error = mean(is.na(B))) %>% 
  filter(prop.error > 0) %>% 
  kable(digits = 3) %>%
  kable_styling("striped")

```

#### Table of Results

In the table, bias, variance, and MSE are all multiplied by a factor of 100. "Neither" means neither nuisance model is correct, "X Only" means only $X|Z$ is correct, etc.

```{r message = F}

# table with simulation results
make.table <- function(param., n., format = "html") {
  
  tbl1 <- sim.out.long %>%
    filter(param == param.,
           n == n.) %>% 
    group_by(q, method, nuis.x, nuis.c) %>% 
    summarise(v1_Bias = 10*mean(B - B.true, na.rm = T),
              v2_ESE = 10*sd(B, na.rm = T),
              v3_ASE = 10*mean(V, na.rm = T),
              v4_Cov = 100*mean(ci.cov, na.rm = T)) %>% 
  gather(key, value, v1_Bias:v4_Cov) %>% 
  unite(Group, q, key) %>% 
  spread(Group, value)
  
  kbl1 <- setNames(tbl1, c("Method", "X|Z", "C|Z",
                   sub(".+_", "", tail(names(tbl1), -3)))) %>% 
    kable(digits = c(0, 0, 0,
                     2, 2, 2, 1,
                     2, 2, 2, 1),
          format = format,
          align = "lccrrrrrrrr",
          booktabs = T) %>%
    kable_styling("striped") %>%
    row_spec(row = 0, bold = TRUE)
    add_header_above(c(" " = 3,
                       "q = 0.4" = 4,
                       "q = 0.8" = 4))
  
  return(kbl1)
  
}

```


```{r message = F}

print("param = 2, n = 200"); make.table(param. = 2, n. = 200)
print("param = 2, n = 800"); make.table(param. = 2, n. = 800)
print("param = 2, n = 8000"); make.table(param. = 2, n = 8000)

```

```{r message = F}

# table with X coefficient only and n = 8000
t2 <- make.table(param. = 2, n. = 8000, format = "latex")

```

```{r message = F}

# table and all parameters
make.table.allparam <- function(n., format = "html") {
  
  tbl1 <- sim.out.long %>%
    filter(n == n.) %>% 
    mutate(Param = ifelse(param == 1, "$\\beta_1$",
                   ifelse(param == 2, "$\\beta_2$",
                   ifelse(param == 3, "$\\beta_3$",
                          "$\\log\\sigma^2$")))) %>% 
    group_by(q, Param, method, nuis.x, nuis.c) %>% 
    summarise(v1_Bias = 10*mean(B - B.true, na.rm = T),
              v2_ESE = 10*sd(B, na.rm = T),
              v3_ASE = 10*mean(V, na.rm = T),
              v4_Cov = 100*mean(ci.cov, na.rm = T)) %>% 
  gather(key, value, v1_Bias:v4_Cov) %>% 
  unite(Group, q, key) %>% 
  spread(Group, value)
  
  kbl1 <- setNames(tbl1, c("Param", "Method", "X|Z", "C|Z",
                   sub(".+_", "", tail(names(tbl1), -4)))) %>% 
    kable(digits = c(0, 0, 0, 0,
                     2, 2, 2, 1,
                     2, 2, 2, 1),
          format = format,
          align = "ccccrrrrrrrr",
          booktabs = T) %>%
    kable_styling("striped") %>%
    row_spec(row = 0, bold = TRUE) %>% 
    add_header_above(c(" " = 4,
                       "q = 0.4" = 4,
                       "q = 0.8" = 4))
  
  return(kbl1)
  
}

```

```{r message = F}

print("n = 200"); make.table.allparam(n. = 200)
print("n = 800"); make.table.allparam(n. = 800)
print("n = 8000"); make.table.allparam(n. = 8000)

# table with n = 8000 and all parameters
t3 <- make.table.allparam(n. = 8000, format = "latex")

```

```{r message = F}

# table with given parameter and all sample sizes
make.table.alln <- function(param., format = "html") {
  
  tbl1 <- sim.out.long %>%
    filter(param == param.) %>% 
    group_by(q, method, nuis.x, nuis.c, n) %>% 
    summarise(v1_Bias = 10*mean(B - B.true, na.rm = T),
              v2_ESE = 10*sd(B, na.rm = T),
              v3_ASE = 10*mean(V, na.rm = T),
              v4_Cov = 100*mean(ci.cov, na.rm = T),
              v5_FTC = 100*mean(is.na(B))) %>% 
  gather(key, value, v1_Bias:v5_FTC) %>% 
  unite(Group, q, key) %>% 
  spread(Group, value)
  
  kbl1 <- setNames(tbl1, c("Method", "X|Z", "C|Z", "n",
                   sub(".+_", "", tail(names(tbl1), -4)))) %>% 
    kable(digits = c(0, 0, 0, 0,
                     2, 2, 2, 1, 1,
                     2, 2, 2, 1, 1),
          format = format,
          align = "lcccrrrrrrrrrr",
          booktabs = T,
          linesep = c("", "", "\\addlinespace")) %>%
    kable_styling("striped") %>%
    row_spec(row = 0, bold = TRUE) %>% 
    add_header_above(c(" " = 4,
                       "q = 0.4" = 5,
                       "q = 0.8" = 5))
  
  return(kbl1)
  
}

```


```{r message = F}

print("param = 2, n = 200"); make.table.alln(param. = 2)
t4 <- make.table.alln(param. = 2, format = "latex")

```

```{r message = F}

# table with beta_1, n = 8000 and median SE
tbl5 <- sim.out.long %>%
  filter(param == 2,
         n == 8000) %>% 
  group_by(q, method, nuis.x, nuis.c) %>% 
  summarise(v1_Bias = 10*mean(B - B.true, na.rm = T),
            v2_ESE = 10*sd(B, na.rm = T),
            v3_ASE = 10*mean(V, na.rm = T),
            v4_MedSE = 10*median(V, na.rm = T),
            v5_Cov = 100*mean(ci.cov, na.rm = T)) %>% 
  gather(key, value, v1_Bias:v5_Cov) %>% 
  unite(Group, q, key) %>% 
  spread(Group, value)

setNames(tbl5, c("Method", "X|Z", "C|Z",
                 sub(".+_", "", tail(names(tbl5), -3)))) %>% 
  kable(digits = c(0, 0, 0,
                   2, 2, 2, 2, 1,
                   2, 2, 2, 2, 1),
        format = "html",
        align = "lccrrrrrrrr",
        booktabs = T) %>%
  kable_styling("striped") %>%
  row_spec(row = 0, bold = TRUE) %>% 
  add_header_above(c(" " = 3,
                   "q = 0.4" = 5,
                   "q = 0.8" = 5))
  
```

#### Plot of Results

```{r}

pal_light <- c('#EE6677', '#228833', '#4477AA', '#CCBB44',
               '#66CCEE','#AA3377', '#BBBBBB')
font.size <- 12

sim.out.long %>%
  filter(param == 2) %>%
  mutate(n = factor(n)) %>% 
  ggplot(aes(y = B,
             x = n,
             fill = method)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = B.true),
             linetype = "dashed",
             linewidth = 0.75) +
  facet_nested(q ~ method + nuis.x + nuis.c,
               scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_text(size = font.size),
        strip.text.x = element_text(size = font.size)) +
  labs(y = "Estimated Coefficient") +
  guides(fill = "none", color = "none") +
  scale_fill_manual(values = pal_light[c(1, 2, 3, 5)])


```

#### Plot of Computation Times

```{r}

sim.out.time <- sim.out %>% 
  select(n, q, starts_with("time")) %>% 
  pivot_longer(starts_with("time"),
               names_to = "proc",
               values_to = "time") %>% 
  mutate(proc = gsub("time.", "", proc)) %>% 
  separate(proc, into = c("proc1", "proc2"), sep = "[.]") %>% 
  mutate(Proc1 = factor(proc1,
                        levels = c("xz", "cz", "est", "var"),
                        labels = c("X|Z Model", "C|Z Model",
                                   "Parameter Estimation",
                                   "Variance Estimation"),
                        ordered = T),
         Proc2 = factor(proc2,
                        levels = c("sp", "ml", "cc", "or",
                                   "param", "nonpar"),
                        labels = c("SPARCC", "MLE",
                                   "Complete Case", "Oracle",
                                   "Parametric", "Nonparametric"),
                        ordered = T))

```

```{r}

# plot times for nuisance model fitting
sim.out.time %>% 
  filter(proc1 %in% c("cz", "xz")) %>% 
  mutate(n = factor(n),
         q = factor(q)) %>% 
  ggplot(aes(x = n,
             y = time)) +
  geom_boxplot(aes(fill = q)) +
  facet_grid(Proc2 ~ Proc1,
             scales = "free") +
  labs(x = "Sample Size (n)",
       y = "Computation Time (Seconds)",
       fill = "Censoring Proportion (q)") +
  theme(legend.position = "bottom")

# plot times for outcome model fitting
sim.out.time %>% 
  filter(proc1 %in% c("est", "var")) %>% 
  mutate(n = factor(n),
         q = factor(q)) %>% 
  ggplot(aes(x = n,
             y = time)) +
  geom_boxplot(aes(fill = q)) +
  facet_grid(Proc1 ~ Proc2,
             scales = "free") +
  labs(x = "Sample Size (n)",
       y = "Computation Time (Seconds)",
       fill = "Censoring Proportion (q)") +
  theme(legend.position = "bottom")

```

#### Table of Computation Times

```{r message = F}

tbl6 <- sim.out.time %>%
  filter(#proc1 %in% c("est", "var"),
         !is.na(time)) %>% 
  group_by(n, q, Proc1, Proc2) %>% 
  summarise(mean = mean(time),
            sd = sd(time)) %>% 
  mutate(time = paste0(round(mean, 1), " (", round(sd, 2), ")")) %>% 
  select(n, q, Proc1, Proc2, mean) %>% 
  gather(key, value, mean) %>% 
  unite(Group, n, q, key) %>% 
  spread(Group, value)

tbl6

kbl6 <- setNames(tbl5, c("Task", "Method",
                 sub(".*_(.+?)_.*", "\\1", tail(names(tbl6), -2)))) %>% 
  kable(digits = c(0, 0,
                   rep(1, 6)),
        format = "latex",
        align = "ccrrrrrr",
        booktabs = T) %>%
  kable_styling("striped") %>%
  row_spec(row = 0, bold = TRUE) %>% 
  add_header_above(c(" " = 2,
                   "n = 200" = 2,
                   "n = 800" = 2,
                   "n = 8000" = 2))

```






