---
title: "Sparcc Simulation 5"
author: "Brian Richardson"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  html_document:
    code_folding: hide
---

```{r message = F, warning = F}

rm(list = ls())
library(dplyr)
library(tidyverse)
library(ggplot2)
library(stringr)
library(ggh4x)
library(scales)
library(ggthemes)
library(kableExtra)

```

## Simulation Setup

The goal of this program is to evaluate the performance of the semiparametric efficient estimator $\widehat{\pmb{\beta}}_{sp}$ using simulations in the following setting:

* $Z \sim \textrm{Bernoulli}(0.5)$,
* $X|Z \sim \textrm{beta}(\alpha_1 + \alpha_2 Z, \alpha_3 + \alpha_4 Z)$,
* $C|Z \sim \textrm{beta}(\gamma_1 + \gamma_2 Z, \gamma_3 + \gamma_4 Z)$,
* $Y|X,Z \sim N(\beta_0 + \beta_1X + \beta_2Z, \sigma^2)$.

The assumption that $Y|X,Z$ follows a normal distribution allows us to use Gaussian quadrature to efficiently solve integrals with respect to $y$. 

In these simulations, we vary the censoring probability $q=\textrm{P}(X>C)$.

We also vary which how nuisance distributions are modeled: $X|Z$ (or $C|Z$) correctly specified as beta conditional on $Z$ with estimated parameters $\pmb{\alpha}, \pmb{\gamma}$, $X|Z$ (or $C|Z$) incorrectly specified as marginal beta, or both $X|Z$ and $C|Z$ nonparametrically estimated.

For each simulation, we generate a data set of size $n$ then compute the complete case estimator $\widehat{\pmb{\beta}}_{cc}$, the oracle estimator $\widehat{\pmb{\beta}}_{or}$, the parametric MLE $\widehat{\pmb{\beta}}_{ml}$, and the semiparametric efficient estimator $\widehat{\pmb{\beta}}_{sp}$.

```{r}

# true beta
B.true <- c(1, 10, 2, log(1))

# load simulation results from each of 10 clusters
sim.out.list <- lapply(
  X = 0:9,
  FUN = function(clust) {
    cbind(clust,
          read.csv(paste0(
            "sim5_data/v3/sd",
            clust, ".csv")))
  })

# combine simulation results into 1 data frame
sim.out <- bind_rows(sim.out.list)

colnames(sim.out) <- gsub(".B", ".", colnames(sim.out))
colnames(sim.out) <- gsub("B2", "trueB2", colnames(sim.out))

# make long data frame
sim.out.long <- sim.out %>% 
  pivot_longer(cols = starts_with(c("B", "V")),
               names_to = "method.param",
               values_to = "estimate") %>% 
  mutate(name = factor(str_sub(method.param, 1, 1)),
         method = factor(str_sub(method.param, 3, 4),
                         levels = c("or", "cc", "ml", "sp"),
                         labels = c("Oracle", "Complete Case",
                                    "Parametric MLE", "Semiparametric")),
         nuis = factor(gsub("[.]", "", str_sub(method.param, 5, -2)),
                            levels = c("", "0", "1", "2",
                                       "00", "01", "10", "11", "22"),
                            labels = c("",
                                       "X|Z Incorrect",
                                       "X|Z Correct",
                                       "X|Z Nonparametic",
                                       "X|Z & C|Z Incorrect",
                                       "X|Z Incorrect & C|Z Correct",
                                       "X|Z Correct & C|Z Incorrect",
                                       "X|Z & C|Z Correct",
                                       "X|Z & C|Z Nonparametric")),
         param = factor(str_sub(method.param, -1, -1)),
         B.true = B.true[param]) %>% 
  dplyr::select(-c(method.param)) %>% 
  group_by(clust, n, q, name, method, nuis, param) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = name,
              values_from = estimate,
              id_cols = c(clust, n, q, method, nuis, param, B.true, id)) %>% 
  mutate(ci.lower = B - qnorm(0.975) * V,
         ci.upper = B + qnorm(0.975) * V,
         ci.cov = B.true >= ci.lower & B.true <= ci.upper)
             
```

## Simulation Results

#### Errors in Simulations

We first look at which simulation settings lead to errors.

```{r message = F}

# check for simulations with errors
sim.out.long %>% 
  filter(param == 1) %>% 
  group_by(q, n, method, nuis) %>% 
  summarize(n.error = sum(is.na(B)),
            prop.error = mean(is.na(B))) %>% 
  filter(prop.error > 0) %>% 
  kable(digits = 3) %>%
  kable_styling("striped")

```

#### Table of Results

In the table, bias, variance, and MSE are all multiplied by a factor of 100. "Neither" means neither nuisance model is correct, "X Only" means only $X|Z$ is correct, etc.

```{r message = F}

make.table <- function(param.) {
  
  tbl1 <- sim.out.long %>%
    filter(param == param.) %>% 
    group_by(q, method, nuis) %>% 
    summarise(v1_Bias = 10*mean(B - B.true, na.rm = T),
              v2_ESE = 10*sd(B, na.rm = T),
              v3_ASE = 10*mean(V, na.rm = T),
              v4_Cov = 100*mean(ci.cov, na.rm = T)) %>% 
  gather(key, value, v1_Bias:v4_Cov) %>% 
  unite(Group, q, key) %>% 
  spread(Group, value)
  
  kbl1 <- setNames(tbl1, c("Method", "Nuisance Model",
                   sub(".+_", "", tail(names(tbl1), -2)))) %>% 
    kable(digits = c(0, 0, 2, 2, 2, 1, 2, 2, 2, 1)) %>%
    kable_styling("striped") %>%
    add_header_above(c(" " = 2,
                       "q = 0.4" = 4,
                       "q = 0.8" = 4))
  
  return(kbl1)
  
}

```

```{r message = F}

make.table(1)
make.table(2)
make.table(3)
make.table(4)

```

